---
title: "Assignment_4"
output: html_document
---
```{r}
library(tidyverse)
```
install.package(latex)

## 3 Sampling with replacement

You can compute this number straightforwardly within R via the choose function choose( ). For example,
if we want to compute the number of different subsets of size 3 from a collection of size 8 we would compute
```{r}
choose(8,3)
```
Suppose we have a bag containing 10 spheres. This includes 3 red spheres and 7 blue spheres.
Let’s suppose that we draw a sphere at random from the bag (all spheres have equal probability of being
drawn). We record its colour and then return the sphere to the bag. This process is repeated 35 times. This
is an example of sampling with replacement since the spheres are replaced after each draw.
Write down a mathematical expression for the probability that z out of the 35 selections were red spheres
(here z ∈ {1, . . . , 35}).
Try doing this with in laTex, making use of the laTex functions binom{}{} and frac{}{}.
Next write an R function called prob_red_spheres() which takes z as an argument and computes the
probability that z out of a total of the 35 balls selected are red.
Test your function as follows
```{r}
rep <- 35

prob_red_spheres <- function(z){
  choose(rep,z)*(0.3^(z))*(0.7^(rep-z))
}

prob_red_spheres(20)
```
$\binom{10}{5}$

Generate a data frame called prob_by_num_reds with two columns num_reds and prob. The num_reds
column should contain numbers 1 through 35 and the prob column should give the associated probability of
selecting that many reds out of a total number of 35 selections.
Display the first 3 rows of your data frame:

Now use the geom_line() function within the ggplot2 library, in conjunction with your data frame to
display a plot of the probability as a function of the number of reds. Your plot should look as follows:

```{r}
num_reds <- seq(1,35)
prob_by_num_reds <- data.frame(num_reds) %>%
  mutate(prob=map_dbl(.x=num_reds,prob_red_spheres))
prob_by_num_reds %>% head(3)

prob_by_num_reds %>% ggplot(aes(num_reds,prob)) +
  geom_line()
```

Next we shall explore the sample() function within R. Let’s suppose we want to simulate a random experiment in which we sample with replacement from a collection of 10 objects, and repeat this process 35 times.
We can do this by calling:

```{r}
sample(10,35,replace=TRUE)
```

Try this out for yourself. The output should be a vector of length 35 consisting entirely of numbers between 1
and 10. Since this is sampling with replacements and the number of samples exceeds the number of elements
there will be repetitions.
Try rerunning the function. You probably get a different sample. This is to be expected, and even desirable,
since the function simulates a random sample. However, the fact that we get a different answer every time
we run the code is problematic from the prespective of reproducibility. To avoid this process we can set a
random seed via the function set.seed(). By doing so we should get the same output every time. Try the
following out for your self:

```{r}
## Setting the random seed just once
set.seed(0)
for(i in 1:5){
print(sample(100,5,replace=FALSE))
  # The result may well differ every time
}
## Resetting the random seed every time
for(i in 1:5){
set.seed(1)
print(sample(100,5,replace=FALSE))
# The result should not change
}

```
We shall now use the sample() to construct a simulation study to explore the probability of selecting z red
balls from a bag of size 10, with 3 red and 7 blue balls, when sampling 35 balls with replacement.
First set a random seed. Then create a data frame called sampling_with_replacement_simulation consisting of a two columns. The first is called trial and contains numbers 1 through 1000. The second is
called sample_balls and corresponds to a random sample of size 35 from a bag of size 10, with replacement.
We can do this as follows:
```{r}
num_trials<-1000 # set the number of trials
set.seed(0) # set the random seed

sampling_with_replacement_simulation<-data.frame(trial=1:num_trials)%>%
  mutate(sample_balls=map(.x=trial,~sample(10,35,replace = TRUE))) %>%
  mutate(num_reds=map_dbl(.x=sample_balls,~sum(.x<=3)))
# generate collection of num_trials simulations
sampling_with_replacement_simulation
```

Now add a new column called num_reds such that, for each row, num_reds contains an integer which gives
the number of items within the sample for that row (the entry in the sample_balls column) which are less
than or equal to three. For example, suppose that some row of the data frame, the sample_balls column
contains the following list:
9, 4, 7, 1, 2, 7, 2, 3, 1, 5, 5, 10, 6, 10, 7, 9, 5, 5, 9, 9,
5, 5, 2, 10, 9, 1, 4, 3, 6, 10, 10, 6, 4, 4, 10
Then the corresponding row of the num_reds column should contain the number 8, since 8 of these values
are less than equal to 3. You may want to use the functions mutate(), map_dbl and sum().
Next we shall add a new column called simulation_count to our existing data frame prob_by_num_reds
which gives the number of times within our simulation we observed the corresponding number of reds. We
can do this as follows:

```{r}
num_reds_in_simulation<-sampling_with_replacement_simulation%>%pull(num_reds)
# we extract a vector corresponding to the number of reds in each trial

prob_by_num_reds<-prob_by_num_reds%>%
  mutate(simulation_count=map_dbl(.x=num_reds,~sum(num_reds_in_simulation==.x)))
# add a column which gives the number of trials with a given number of reds
prob_by_num_reds %>% head(5)
```
Next we add a column called expected_count corresponding to the expected number of observed reds in
1000 based upon your probability formula.

```{r}
prob_by_num_reds<-prob_by_num_reds%>%
  mutate(expected_count=num_trials*prob)
prob_by_num_reds %>% head(5)
# add a column which gives the expected number of reds
# based on the probability formula
```
Finally, create a plot which compares the results of your simulation with the expected count based on your
probability formula. The concept of expectation will be discussed in Lecture 11.
Your result should look something like the plot below. Of course, since this is a random simulation, your
result may well look slightly different.

```{r}
prob_by_num_reds%>%
rename(Simulation=simulation_count,Expected=expected_count)%>%
pivot_longer(cols=c("Simulation","Expected"),
names_to="Type",values_to="count")%>%
ggplot(aes(num_reds,count)) +
geom_line(aes(linetype=Type, color=Type)) +
scale_linetype_manual(values = c("solid", "dashed"))+
theme_bw()+
xlab("Number of reds")+
ylab("Count")
```
## 4 Sampling without replacement
This question is more challenging. However, you should aim to complete at least the simulation component
using ideas from the previous question.
Let’s suppose we have large bag containing 100 spheres. There are 50 red spheres, 30 blue spheres and 20
green spheres. Suppose that we sample 10 spheres from the bag without replacement.
What is the probability that one or more colours is missing from your selection?
First aim to answer this question via a simulation study using ideas from the previous question.
You may want to use the following steps:
1. First set a random seed;
2. Next set a number of trials, a number of reds, a number of blues, a number of greens and a sample
size;
3. Now use a combination of the functions sample(), mutate() and map() to generate your samples.
Here you are creating sample of size 10 from a collection of 100 balls - the sampling is done without
replacement;
4. Now compute the number of “reds”, “greens” and “blues” in your sample using the map_dbl() and
mutate() functions;
5. Compute the minimum of the three counts using the pmin() function. When this minimum is zero,
then one of the three colours is missing. It is recommended that you look up the difference between
pmin() and min() here;
6. Compute the proportion of rows for which the minimum number of the three counts is zero.
Try this initially with a small number of simulations. Increase your number of simulations to about a
relatively large number to get a more accurate answer, once everything seems to be working well.
The next part of the question is more challenging and may be omitted if you are short on time.
(*) Once you have a simulation based answer you can try and use “combinations” with n
k

to compute the
probability directly. First aim and compute the number of subsets of size 10 from 100 which either entirely
miss out one of the subsets Reds = {1, . . . , 50}, Blues = {51, . . . , 80}, Greens = {81, . . . , 100}. Be careful
not to double count some of these subsets! Once you have computed all such subsets combine with the
formula for the total number of subsets of size 10 from a set of 100, to compute the probability of missing a
colour
```{r}
num_trials<-1000 # set the number of trials
set.seed(0) # set the random seed

sampling_without_replacement_simulation<-data.frame(trial=1:num_trials)%>%
  mutate(sample_balls=map(.x=trial,~sample(100,10,replace = F))) %>%
  mutate(num_reds=map_dbl(.x=sample_balls,~sum(.x<=50))) %>%
  mutate(num_blues=map_dbl(.x=sample_balls,~sum(.x>50 & .x<=80))) %>%
  mutate(num_greens=map_dbl(.x=sample_balls,~sum(.x>80 & .x<=100))) %>%
  mutate(minimum=pmin(num_reds,num_blues,num_greens)) 
sampling_without_replacement_simulation %>% head(5)

zeros <- nrow(sampling_without_replacement_simulation[sampling_without_replacement_simulation$minimum == "0",])
proportion_zeros <- zeros/nrow(sampling_without_replacement_simulation)
proportion_zeros
# generate collection of num_trials simulations

z <- 10
rep <- 10

all_reds <- choose(num_trials,z)*(0.5^z)*(0.3^(1-z))*(0.2^(1-z))
no_reds <- 1-all_reds
no_reds

prob_red_spheres <- function(z){
  choose(rep,z)*(0.5^z)*(0.3^(1-z))*(0.2^(1-z))
}

prob_red_spheres(10)
```

```{r}


```

```{r}

```

```{r}

```


















