  ---
title: "Assignment_2"
output: html_document
---
---
title: "Assignment_2"
output: html_document
---

install.packages("tidyverse")
library(tidyverse)
install.packages("Stat2Data")
library(Stat2Data)
data("Hawks")
hawksSmall<-drop_na(select(Hawks,Age,Day,Month,Year,CaptureTime,Species,Wing,Weight,Tail))

```{r}
# Check how many rows and columns hawkssmall using the dim() function.
dim(hawksSmall)
# Display the top 5 rows of the hawkssmall data frame
head(hawksSmall)
```

# 1.3 Generate a histogram
```{r}
hawk_weight <- ggplot(data=hawksSmall, aes(x=Weight))+xlab("Weight (gm)")
hawk_weight+geom_histogram(binwidth=100)+ylab("Count")
```

# 1.4 Generate a density plot
```{r}
hawk_tail <- ggplot(data=hawksSmall, aes(x=Tail))+xlab("Tail (mm)")
hawk_tail+ geom_density(adjust=0.5)+ylab("Density")
# adjust at 0.5 shows a Trimodal density plot while at 1 gives a Bimodal density plot

hawk_tail2 <- ggplot(data=hawksSmall, aes(x=Tail,color=Species))+
  geom_density(adjust=1)+theme_bw()+xlab("Tail (mm)")+ylab("Density")
hawk_tail2

hawk_tail3 <- ggplot(data=hawksSmall, aes(x=Tail,y=Species,fill=Species))+
  geom_violin()+theme_bw()+xlab("Tail (mm)")+ylab("Species")
hawk_tail3
```
# 1.5 Scatter plots
# two aes: color and shape

```{r}
hawk_tail4 <- ggplot(data=hawksSmall, aes(x=Tail,y=Weight))+
  geom_point(aes(color=Species,shape=Species))
hawk_tail4
```

# 1.6 Trend lines and facet wraps
```{r}
hawk_tail5 <- ggplot(data=hawksSmall, aes(x=Tail,y=Weight))+
  geom_point(aes(color=Species)) +
  facet_wrap(~Species,scales="free")+
  geom_smooth(method="lm")+
  theme_bw()+
  xlab("Tail (mm)")+
  ylab("Weight (gm)")
hawk_tail5
```

# 2 Data wrangling

#2.1 Select and filter functions
```{r}
head(Hawks)
hSF <- Hawks %>% 
  filter(Species=="RT",Weight>c(1000)) %>%
  select("Wing","Weight","Tail")
hSF
dim(hSF)
# hSF contains 3 variables with 387 cases
```

# 2.2 The arrange function
```{r}
hSF %>% arrange(Wing)
```
# 2.3 Join and rename functions
```{r}
species_code <- unique(Hawks$Species)
species_name_full <- c("Red-tailed","Cooper's","Sharp-shinned")
full_name_df <- data.frame(species_code,species_name_full)
full_name_df

HawksFullName <- Hawks %>% 
  rename(species_code = Species) %>%
  left_join(full_name_df) %>%
  rename(Species = species_name_full) %>%
  select(-species_code)
HawksFullName


head(HawksFullName %>% select("Species","Wing","Weight"))
```
# 2.4 The mutate function
```{r}
hawksWithBMI <- Hawks %>%
  mutate(bird_BMI = 1000*Weight/Wing^(2)) %>%
  select("Species","bird_BMI") %>%
  arrange(desc(bird_BMI)) %>%
  filter(bird_BMI<c(100)) %>%
  ggplot(aes(x=bird_BMI,y=Species,fill=Species)) +
  geom_violin() +
  scale_x_continuous()+
  theme_bw() +
  xlab("Bird_BMI") +
  ylab("Species")

hawksWithBMI
```
# 2.5 Summarize and group-by functions
```{r}
HawksFullName %>%
  group_by(Species) %>%
  summarize(n_rows=n(), 
            avg_wing_span = mean(Wing, na.rm=T),
            median_wing_span = median(Wing, na.rm=T),
            trimmed_mean_avg_wing_span = mean(Wing, na.rm=1,trim=0.1),
            wing_tail_ratio = mean(Wing/Tail,na.rm=T)
  )

HawksFullName %>%
  group_by(Species) %>%
  select("Wing", "Weight", "Culmen", "Hallux", "Tail", "StandardTail", "Tarsus", "Crop") %>%
  summarize(across(everything(),~sum(is.na(.x)))) 
```
# 3 Exploratory data analysis

# 3.1 Combining location estimators with the summarise function
```{r}
Hawks %>%
  summarize(Wing_mean = mean(Wing, na.rm=T),
            Wing_t_mean = mean(Wing, na.rm=1, trim=0.1),
            Wing_med = median(Wing, na.rm=T),
            Weight_mean = mean(Weight, na.rm=T),
            Weight_t_mean = mean(Weight, na.rm=1, trim=0.1),
            Weight_med = median(Weight, na.rm=T)
  )

Hawks %>%
  group_by(Species) %>%
  summarize(Wing_mean = mean(Wing, na.rm=T),
            Wing_t_mean = mean(Wing, na.rm=1, trim=0.1),
            Wing_med = median(Wing, na.rm=T),
            Weight_mean = mean(Weight, na.rm=T),
            Weight_t_mean = mean(Weight, na.rm=1, trim=0.1),
            Weight_med = median(Weight, na.rm=T)
  )

  #group_by(Species) %>%

```
# 3.2 Location and dispersion estimatiors under linear transformations

# 3.3 Robustness of location estimators
```{r}
hal<-Hawks$Hallux # Extract the vector of hallux lengths
hal<-hal[!is.na(hal)] # Remove any nans

outlier_val<-100
num_outliers<-10
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))

mean(hal)
mean(corrupted_hal)

num_outliers_vect<-seq(0,1000)
means_vect<-c()
for(num_outliers in num_outliers_vect){
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
means_vect<-c(means_vect,mean(corrupted_hal))
medians_vect<-c(medians_vect,median(corrupted_hal))
t_means_vect<-c(t_means_vect,mean(corrupted_hal,trim=0.1))
}
```

```{r}
```









