---
title: "Assignment_3"
output: html_document
---
```{r}
library(tidyverse)
```
# 2 Tidy data and iteration

## 2.1 Missing data and iteration
```{r}
impute_by_median <- function(x){
  med<-median(x,na.rm=1) 
  impute_f <- function(y){
    if(is.na(y)){
      return(med)
    }else{
      return(y)
    }
  }
  return(map_dbl(x,impute_f))
}
v<-c(1,2,NA,4)
impute_by_median(v)
```

Next generate a data frame with two variables x and y. For our first variable x we have a sequence
(x1, x2, . . . , xn) where x1 = 0, xn = 10 and for each i = 1, . . . , n − 1, xi+1 = xi + 0.1. For our second
variable y we set yi = 5 × xi + 1 for i = 1, . . . , n. Generate data of this form and place within a data frame
called df_xy.

```{r}
x <- seq(0,10,0.1)
y<-5*x+1
df_xy <- data.frame(x,y)
df_xy %>% head(5)
```

The map2() function is similar to the map() function but iterates over two variables in parallel rather
than one. You can learn more here https://purrr.tidyverse.org/reference/map2.html. The following simple
example shows you how map2_dbl() can be combined with the mutate() function.

```{r}
df_xy%>%
mutate(z=map2_dbl(x,y,~.x+.y))%>%
head(5)
```

First create a function sometimes_missing with two variables index and value. The function should return
NA if index is divisible by 5 and returns value otherwise.

```{r}
sometimes_missing <- function(index,value){
  if (index%%5 == 0){
    return(NA)
    }
  else{
    return(value)
    }
}
sometimes_missing(14,25)
sometimes_missing(15,25)
```

Next generate a new data frame called df_xy_missing with two variables x and y, but some missing data.
For the first variable x we have a sequence (x1, · · · , xn), which is precisely the same as with df_xy. For
the second variable y we have a sequence (˜y1, · · · , y˜n) where y˜i = NA if i is divisible by 5 and y˜i = yi for
i not divisible by 5. To generate the dataframe d_xy_missing you may want to make use of the functions
row_number(), map2_dbl(), mutate() as well as sometimes_missing().

```{r}

df_xy_missing <- df_xy %>%
  mutate(y=map2_dbl(.x=row_number(),.y=y,~sometimes_missing(.x,.y)))
df_xy_missing
```

Create a new data frame df_xy_imputed with two variables x and y. For the first variable x we have a
sequence (x1, · · · , xn), which is precisely the same as with df_xy. For the second variable y we have a
sequence (y
′
1
, · · · , y′
n
) which is formed from (˜y1, · · · , y˜n) by imputing any missing values with the median.
To generate df_xy_imputed from “‘df_xy_missing by applying a combination of the functions mutate and
impute_by_median().

```{r}
df_xy_imputed <- df_xy_missing %>%
  mutate(y=impute_by_median(y))
df_xy_imputed
```

Combine the dataframes df_xy, df_xy_missing and df_xy_impute within a single dataframe called
df_combined, along with an additional column indicating the source of the data.

```{r}
nrow(df_xy_missing)
nrow(df_xy_imputed)
df_xy<-df_xy%>%
mutate(source="original")
df_xy_missing<-df_xy_missing%>%
mutate(source="corrupted")
df_xy_imputed<-df_xy_imputed%>%
mutate(source="imputed")
df_combined<-rbind(df_xy,df_xy_missing,df_xy_imputed)
df_combined
```

Plot the original data, the corrupted data and the imputed data together together with a trend line for each
sample.

```{r}
ggplot(df_combined,aes(x=x,y=y,color=source))+geom_point()+
facet_wrap(~source)+geom_smooth(method="lm")

```
## 2.2 Tidying data with pivot functions

In this task you will read in data from a spreadsheet and apply some data wrangling tasks to tidy that data.
First download the excel spreadsheet entitled “HockeyLeague.xlsx”. The excel file contains two spreadsheets - one with the wins for each team and one with the losses for each team. To read this spreadsheet
into R we shall make use of the readxl library. You may need to install the library:

```{r}
install.packages("readxl")
```

The following code shows how to read in a sheet within an excel file as a data frame. You will need to edit
the folder_path variable to be the directory which contains your copy of the spreadsheet.

```{r}
library(readxl) # load the readxl library
library(tidyverse)
folder_path<-"C:/Users//ov21312/OneDrive - University of Bristol/Documents/Rproject/UoBRProject/" # set this to the name of the
# directory containing "HockeyLeague.xlsx"
file_name<-"HockeyLeague.xlsx" # set the file name
file_path<-paste(folder_path,file_name,sep="") # create the file_path
wins_data_frame<-read_excel(file_path,sheet="Wins") # read of a sheet from an xl file
#Inspect the first 3 rows of the first five columns:
wins_data_frame %>%
select(1:5)%>%
head(3)
```
A cell value of the form “a of b” means that a games were won out of a total of b for that season. For
example, the element for the “Ducks” row of the “1990” column is “30 of 50” meaning that 30 out of 50
games were won that season.
Is this tidy data?
Now apply your data wrangling skills to transform the “wins_data_frame” data frame object into a data
frame called “wins_tidy” which contains the same information but has just four columns entitled “Team”,
“Year”, “Wins”, “Total”. The “Team” column should contain the team name, the “Year” column should
contain the year, the “Wins” column should contain the number of wins for that season and the “Total”
column the total number of games for that season. The first column should be of character type and
the remaining columns should be of integer type. You can do this by combining the following functions:
rename(), pivot_longer(), mutate() and separate().
You can check the shape of your data frame and the first five rows as follows:

```{r}
wins_data_frame
wins_tidy <- wins_data_frame %>%
  rename(Team = ...1) %>%
  pivot_longer(!Team,names_to="Year",values_to="Wins") %>%
  separate(Wins,into=c("Wins","Total"),sep=" of ", convert=T) %>%
  mutate(Year= as.integer(Year))

wins_tidy%>% dim() # check the dimensions
wins_tidy%>%head(5) # inspect the top 5 rows
```

The “HockeyLeague.xlsx” also contains a sheet with the losses for each team by season. Apply a similar
procedure to read the data from this sheet and transform that data into a dataframe called “losses_tidy”
with four columns: “Team”, “Year”, “Losses”, “Total” which are similar to thos in the “wins_tidy” data
frame except for the “Losses” column gives the number of losses for a given season and team, rather than
the number of losses.
You may notice that the number of wins plus the number of losses for a given team, in a given year does not
add up to the total. This is because some of the games are neither wins nor losses but draws. That is, for a
given year the number of draws is equal to the total number of games minus the sum of the wins and losses.
Now combine your two data frames, “wins_tidy” and “losses_tidy”, into a single data frame entitled
“hockey_df” which has 248 rows and 9 columns: A “Team” column which gives the name of the team
as a character, the “Year” column which gives the season year, the “Wins” column which gives the number
of wins for that team in the given year, the “Losses” column which gives the number of losses for that team
in the given year and the “Draws” column which gives the number of draws for that team in the given year,
the “Wins_rt” which gives the wins as a proportion of the total number of games (ie. Wins/Total) and
similarly the “Losses_rt” and the “Draws_rt” which gives the losses and draws as a proportion of the total,
respectively. To do this you can make use of the mutate() function. You may also want to utilise the
across() function for a slightly neater solution.

```{r}
losses_data_frame<-read_excel(file_path,sheet="Losses") # read of a sheet from an xl file
#Inspect the first 3 rows of the first five columns:
losses_data_frame %>%
select(1:5)%>%
head(3)

losses_data_frame
losses_tidy <- losses_data_frame %>%
  rename(Team = ...1) %>%
  pivot_longer(!Team,names_to="Year",values_to="Losses" ) %>%
  separate 

losses_tidy

losses_tidy%>% dim() # check the dimensions
losses_tidy%>%head(5) # inspect the top 5 rows
```

```{r}
hockey_df <- inner_join(wins_tidy,losses_tidy) %>%
  mutate(Draws=Total-(Wins+Losses))

%>%
  
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```



















