---
title: "Summative Assessment"
author: "Avekan"
date: "12/11/2021"
output: html_document
---

```{r include = FALSE}
# Load required libraries
# install.packages("plyr")
library(tidyverse)
library(stringr)
```

# Section A

Data wrangling task with some finance data.

## A.1
Load finance data and create data frame "finance_data_original"
```{r}
finance_data_original <- read.csv("./finance_data_EMATM0061.csv",fill=1)
dim(finance_data_original) 
```
The data frame consist of 1051 rows and 30 columns

## A.2
Generate subset data frame "finance_data"
```{r}
finance_data <- finance_data_original %>%
  rename("education_expenditure"="Details.Education.Education.Total",
         "health_expenditure"="Details.Health.Health.Total.Expenditure",
         "transport_expenditure"="Details.Transportation.Highways.Highways.Total.Expenditure",
         "totals_revenue"="Totals.Revenue",
         "totals_expenditure"="Totals.Expenditure") %>%
  select(state_year_code,education_expenditure,health_expenditure,transport_expenditure,totals_revenue,totals_expenditure)

# Display subset of "finance_data" data frame consisting of the first five rows and first three columns
finance_data %>% select(state_year_code,education_expenditure,health_expenditure) %>%
  head(5) 
```

## A.3
Create a new column within "finance_data" data frame called "totals_saving"
```{r}
finance_data <- finance_data %>%
  mutate(totals_savings=totals_revenue-totals_expenditure)

# Display subset of "finance_data" data frame consisting of the first three rows and the four columns
finance_data %>% select(state_year_code, totals_revenue, totals_expenditure ,totals_savings) %>%
  head(3)
```

## A.4
Divide the “state_year_code” column into two separate columns, a “state” column and a “year” column
```{r}
finance_data <- finance_data %>%
  separate(state_year_code, into=c("state","year"), sep="__", convert=T) %>%
  mutate(state=str_to_title(state))

# Display subset of "finance_data" data frame consisting of the first three rows and the five columns
finance_data %>% select(state, year, totals_revenue, totals_expenditure, totals_savings) %>%
  head(3)
```

## A.5
Generate a plot which displays the total revenue (“total_revenue”) as function of the year (“year”) for the following
four states: Louisiana, Montana, Mississippi and Kentucky.
```{r}
finance_data %>%
  filter(state=="Louisiana"|state=="Montana"|state=="Mississippi"|state=="Kentucky") %>% #filters out the four states
  ggplot(aes(x=year,y=totals_revenue/1e+06))+
  geom_smooth(aes(color=state,linetype=state)) + #create geom smooth plot
  theme_bw()+
  labs(y="Revenue (millions $)", x="Year")
```

## A.6
Create a function called get_decade() which takes as input a number and rounds that number down to the
nearest multiple of 10. For example, the numbers 20, 21, 22, . . . , 29 would all be mapped to the output 20.

Use your get_decade() function to add a new column to the “finance_data” data frame called “decade”
which should give the decade corresponding to the year column. For example, the decade of the years
1990,1991,. . . ,1998,1999 is 1990.
```{r}
#Use modulus to extract rightmost digit and subtract from original year to floor year
get_decade <- function(x){ 
  return(x-x%%10)
}

finance_data <- finance_data %>%
  mutate(decade=map_dbl(.x=year,.f=~get_decade(.x))) 

finance_data %>%
  group_by(state) %>%
  filter(decade==c(2000)) %>% #Chooses the decade starting 2000
  select(state, totals_savings) %>% 
  summarise(across(where(is.numeric),~mean(.x,na.rm=T))) %>% #Calculate average savings across decade
  rename("mean_savings_across_decade"=totals_savings) %>%
  arrange(desc(mean_savings_across_decade)) %>% #Arrange in descending order
  head(3)
```
Display of the top 3 total savings countries in descending order

## A.7
Next generate a summary data frame from the “finance_data” data frame called “alaska_summary” 
```{r}
alaska_summary <- finance_data %>%
  group_by(decade) %>%
  filter((decade=="1990"|decade=="2000"|decade=="2010")&state=="Alaska") %>%
  select(decade,education_expenditure,health_expenditure,transport_expenditure) %>%
  summarise(across(ends_with(c("expenditure")), list(md=median,mn=mean), na.rm=T, .names="{substring(.col,1,2)}_{.fn}"))

# Display of "alaska_summary" data frame
alaska_summary
```


## A.8
Create a function called impute_by_median which takes as input a vector numerical values, which may include
some “NA”s, and replaces any missing values (“NA”s) with the median over the vector.
```{r}
impute_by_median <- function(x){
  med<-median(x,na.rm=1) 
  impute_f <- function(y){
    if(is.na(y)){
      return(med)
    }else{
      return(y)
    }
  }
  return(map_dbl(x,impute_f))
}
```

Generate a subset of your “finance_data” data frame called “idaho_2000” 
```{r}
idaho_2000 <- finance_data %>%
  filter(decade==c(2000),state==c("Idaho")) %>%
  select(-state,-decade)
idaho_2000
```

Apply “impute_by_median” function to create a new data frame called “idaho_2000_imputed”
```{r}
idaho_2000_imputed <- idaho_2000 %>%
  mutate(education_expenditure=impute_by_median(education_expenditure),
         health_expenditure=impute_by_median(health_expenditure),
         transport_expenditure=impute_by_median(transport_expenditure),
         totals_revenue=impute_by_median(totals_revenue),
         totals_expenditure=impute_by_median(totals_expenditure),
         totals_savings=impute_by_median(totals_savings))

# Display subset of "idaho_2000_imputed" data frame consisting of the first five rows and the four columns
idaho_2000_imputed %>%
  select(year,health_expenditure,education_expenditure,totals_savings) %>%
  head(5)

```

# Section B

## B.1
### (a)
Write a function called c_prob_person_given_alarm which gives ϕ as a function of p0,p1 and q.
```{r}
c_prob_person_given_alarm <- function(p_0,p_1,q){
  q*p_1/(p_1*q+p_0*(1-q))
}
```

### (b)
Consider a setting in which p0 = 0.05, p1 = 0.95 and q = 0.1. In this case, what is ϕ?
```{r}
p_0 = 0.05
p_1 = 0.95
q = 0.1
c_prob_person_given_alarm(p_0,p_1,q)
```
### (c)
```{r}
q = seq(0.1,1,0.01)

vary_q = data.frame(q) %>%
  mutate(phi=map_dbl(.x=q,.f=~c_prob_person_given_alarm(p_0,p_1,.x))) %>%
  ggplot(aes(x=q,y=phi)) +
  geom_line()
vary_q
```
## B.2

```{r}
set.seed(0)
n=1000

sample_X_0125 <- function(alpha,beta,gamma,n){
  sample_X<-data.frame(U=runif(n))%>%
  mutate(X=case_when(
  (0<=U)&(U<alpha)~1,
  (alpha<=U)&(U<(alpha+beta))~2,
  ((alpha+beta)<=U)&(U<=(1-alpha-beta))~5,
  ((alpha+beta+gamma)<=U)&(U<=(1))~0)) %>%
  pull(X) %>%
  replace_na(0)
  return(sample_X) 
}

alpha = 0.1
sample_X_0125(0.1,0.2,0.3,10000) %>%
  sum(x)/10000

var(x)
```

```{r}
set.seed(0)
num_trials_per_sample_size<-10000
n <- 100
alpha <- 0.1
beta <- 0.2
gamma <- 0.3

simulation_df<-crossing(trial=seq(num_trials_per_sample_size),
                        sample_size=n) %>%
  
  mutate(simulation=pmap(.l=list(trial,sample_size),
                       .f=~sample_X_0125(alpha,beta,gamma,.y)))

simulation_df
```





































































