---
title: "Assignment_5"
output: html_document
---

```{r}
library(tidyverse)
```

# 2 Simulating data with the uniform distribution


```{r}
set.seed(0)
n<-1000
sample_X<-data.frame(U=runif(n))%>%
  mutate(X=case_when(
  (0<=U)&(U<0.25)~1,
  (0.25<=U)&(U<0.5)~5,
  (0.5<=U)&(U<=1)~0))%>%
  pull(X)

sample_X_015 <- function(alpha,beta,n){
  sample_X<-data.frame(U=runif(n))%>%
  mutate(X=case_when(
  (0<=U)&(U<alpha)~1,
  (alpha<=U)&(U<(alpha+beta))~5,
  ((alpha+beta)<=U)&(U<=(1-alpha-beta))~0))%>%
  pull(X) %>%
  replace_na(0)
  return(sample_X) 
}

sample_X_015(0.5,0.1,10000) %>% 
  sum(x)/10000

var(x)


```

```{r}
alpha <- 0.1
beta <- seq(0,0.9,0.01)
n <- 100
set.seed(0)

beta_vary<-data.frame(beta=beta) %>%
  mutate(sample = map(.x=beta,~sample_X_015(alpha,.x,n))) %>%
  mutate(sample_avg = map_dbl(.x=sample,~mean(.x))) %>% 
  ggplot(aes(x=beta,y=sample_avg))+
  geom_line()
  
beta_vary
```
# 3 The Gaussian distribution

Write out the probability density function of a Gaussian random variable with mean µ and standard deviation
σ > 0.
Use the help function to look up the following four functions: dnorm(), pnorm(), qnorm() and rnorm().
Generate a plot which displays the probability density function for three Gaussian distributions:

```{r}
x <- seq(-4,6,0.1)
normal_densities_by_x<-data.frame(x=x,density=dnorm(x,mean=1,sd=sqrt(1)),var=1)%>%
rbind(data.frame(x=x,density=dnorm(x,mean=1,sd=sqrt(2)),var=2))%>%
rbind(data.frame(x=x,density=dnorm(x,mean=1,sd=sqrt(3)),var=3))

normal_densities_by_x%>%
  ggplot(aes(x,y=density,color=as.character(var),linetype=as.character(var)))+
  geom_line()+
  theme_bw()+
  labs(color="Variance",linetype="Variance",x="x",y="Density")


normal_densities_by_x

```

Generate a corresponding plot for the cumulative distribution function for three Gaussian distributions

```{r}
q1 <- data.frame(x=x,cdf=pnorm(x,mean=1,sd=sqrt(1)),var=1)
q2 <- data.frame(x=x,cdf=pnorm(x,mean=1,sd=sqrt(2)),var=2)
q3 <- data.frame(x=x,cdf=pnorm(x,mean=1,sd=sqrt(3)),var=3)

normal_cdf <- q1 %>% rbind(q2) %>% rbind(q3) %>%
  ggplot(aes(x,y=cdf,color=as.character(var),linetype=as.character(var)))+
  geom_line()+
  theme_bw()+
  labs(color="Variance",linetype="Variance",x="x",y="Cumulative distribution function")

normal_cdf
```
Next generate a plot for the quantile function for the same three Gaussian distributions. Describe the relationship
between the quantile function and the cumulative distribution function.


```{r}
prob <- seq(0,1,0.01)

q1 <- data.frame(p=prob,q=qnorm(prob,mean=1,sd=sqrt(1)),var=1)
q2 <- data.frame(p=prob,q=qnorm(prob,mean=1,sd=sqrt(2)),var=2)
q3 <- data.frame(p=prob,q=qnorm(prob,mean=1,sd=sqrt(3)),var=3)
normal_cdf <- q1 %>% rbind(q2) %>% rbind(q3)

ggplot(normal_cdf,aes(x=p,y=q,color=as.character(var),linetype=as.character(var)))+
geom_line()+
theme_bw()+
labs(color="Variance",linetype="Variance",y="Quantile",x="Probability")
```
Now use rnorm() generate a random independent and identically distributed sequence Z1, · · · , Zn ∼ N (0, 1) so
that each Zi ∼ N (0, 1) has standard Gaussian distribution with n = 100. Make sure your code is reproducible
by using the set.seed() function. Store your random sample in a vector called “standardGaussianSample”.

```{r}
set.seed(0)
n=100
standardGaussianSample <- rnorm(n,0,1)
mean1Var3GaussianSampleA <- sqrt(3)*standardGaussianSample+1

set.seed(0)
mean1Var3GaussianSampleB <- rnorm(n,1,sqrt(3))
all.equal(mean1Var3GaussianSampleA,mean1Var3GaussianSampleB)

colors<-c("Population density"="red", "Sample kernel density"="blue",
"Population mean"="green", "Sample mean"="violet")
linetypes<-c("Population density"="solid", "Sample kernel density"="dashed",
"Population mean"="solid", "Sample mean"="dashed")

ggplot()+labs(x="x",y="Density")+theme_bw()+
geom_line(data=(normal_densities_by_x%>%filter(var==3)),
aes(x,y=density,color="Population density"))+
# create plot of theoretical density
geom_density(data=data.frame(x=mean1Var3GaussianSampleA),
aes(x=x,color="Sample kernel density",
linetype="Sample kernel density"))+
# add in kernel density plot from real sample
geom_vline(aes(xintercept=1,color="Population mean",
linetype="Population mean"))+
geom_vline(aes(xintercept=mean(mean1Var3GaussianSampleA),
color="Sample mean",linetype="Sample mean"))+
scale_color_manual(name = "Legend", values=colors)+
scale_linetype_manual(name="Legend", values=linetypes)
```
# 4 The Binomial distribution and the central limit theorem
Consider the case where n = 50 and p = 7/10. Use the dbinom() to generate a dataframe called binom_df with
two columns - x and pmf. The first column contains the numbers {0, 1, . . . , 50} inclusive. The second column
gives the corresponding value of the probability mass function pZ(x) = P(Z = x) with Z ∼ Binom(50, 7/10).
Use the head() function to observe the first 3 rows as your data frame. The result should look as follows:

```{r}
n=50
p=0.7
x=seq(0,50,1)
binom_df <- data.frame(x=x) %>% 
  mutate(pmf=map_dbl(.x=x,~dbinom(x=.x,size=n,prob=p)))

binom_df%>% head(3)
```

```{r}
x<- seq(0,50,0.01)
mu <- n*p
sigma <- n*p*(1-p)

norm_df <- data.frame(x=x) %>%
  mutate(pdf=map_dbl(.x=x,~dnorm(x=.x,mu,sqrt(sigma))))

norm_df %>% head(3)

```
```{r}
colors<-c("Gaussian pdf"="red", "Binomial pmf"="blue")
fill<-c("Gaussian pdf"="white", "Binomial pmf"="white")
ggplot()+labs(x="x",y="Probability")+theme_bw()+
geom_line(data=norm_df,
aes(x,y=pdf,color="Gaussian pdf"),size=2)+
# create plot of Gaussian density
geom_col(data=binom_df,
aes(x=x,y=pmf,color="Binomial pmf",fill="Binomial pmf"))+
scale_color_manual(name = "", values=colors)+
scale_fill_manual(name = "", values=fill)+
xlim(c(20,50))
```
```{r}

```
```{r}

```

