---
title: "Assignment_8"
output: html_document
---
```{r echo=FALSE}
library(tidyverse)
```

# 1 Obstacles to valid scientific inference
1. Measurement error -  difference between the measured value of a quantity and its true value. Eg, rounding errors dude to computational constraints.

2. Selection bias - when the data included in the analysis is mis-representing the underlying population of interest. Different forms of selection bias are not mutally exclusive.

3. Confounding variables - when a third variable (the confounding variable) has a causal effect upon two variables.

# 2 An unpaired t test

```{r}
library(palmerpenguins)
peng_AC<-penguins%>%
drop_na(species,body_mass_g)%>%
filter(species!="Gentoo")
peng_AC 
cat <- unique(peng_AC$species)
cat[1]
```

```{r}

t_test_function <- function(data,val_col,group_col){
  cat <- data%>%pull(group_col)
  categories <- unique(cat) 
  mean_0 <- data%>%filter(!!as.symbol(group_col)==categories[1])%>%pull(val_col)%>%mean()
  mean_1 <- data%>%filter(!!as.symbol(group_col)==categories[2])%>%pull(val_col)%>%mean()
  sd_0 <- data%>%filter(!!as.symbol(group_col)==categories[1])%>%pull(val_col)%>%sd()
  sd_1 <- data%>%filter(!!as.symbol(group_col)==categories[2])%>%pull(val_col)%>%sd()
  n_0 <- data%>%filter(!!as.symbol(group_col)==categories[1])%>%nrow()
  n_1 <- data%>%filter(!!as.symbol(group_col)==categories[2])%>%nrow()

  sd_combined <- sqrt(((n_0-1)*sd_0^2+(n_1-1)*sd_1^2)/(n_0+n_1-2))
  t_statistic <- (mean_0-mean_1)/(sd_combined*sqrt(1/n_0+1/n_1))
  
  p_value <-2*(1-pt(abs(t_statistic),df=n_0+n_1-2))
  dof=n_0+n_1-2
  output <- data.frame(t_statistic,dof,p_value)
  return(output)
}

t_test_function(data=peng_AC,val_col="body_mass_g",group_col="species")

```
```{r}
t_test_function <- function(data,val_col,group_col,var_equal){
  cat <- data%>%pull(group_col)
  categories <- unique(cat) 
  mean_0 <- data%>%filter(!!as.symbol(group_col)==categories[1])%>%pull(val_col)%>%mean()
  mean_1 <- data%>%filter(!!as.symbol(group_col)==categories[2])%>%pull(val_col)%>%mean()
  sd_0 <- data%>%filter(!!as.symbol(group_col)==categories[1])%>%pull(val_col)%>%sd()
  sd_1 <- data%>%filter(!!as.symbol(group_col)==categories[2])%>%pull(val_col)%>%sd()
  n_0 <- data%>%filter(!!as.symbol(group_col)==categories[1])%>%nrow()
  n_1 <- data%>%filter(!!as.symbol(group_col)==categories[2])%>%nrow()
  
  if (var_equal==TRUE){
    sd_combined <- sqrt(((n_0-1)*sd_0^2+(n_1-1)*sd_1^2)/(n_0+n_1-2))
    t_statistic <- (mean_0-mean_1)/(sd_combined*sqrt(1/n_0+1/n_1))
    dof=n_0+n_1-2
    p_value <-2*(1-pt(abs(t_statistic),df=dof))
    
  } else{
    
    t_statistic<-(mean_0-mean_1)/sqrt((sd_0^2/n_0)+(sd_1^2/n_1))
    dof <- (sd_0^2/n_0+sd_1^2/n_1)^2/((sd_0^2/n_0)^2/(n_0-1)+(sd_1^2/n_1)^2/(n_1-1))
    p_value <-2*(1-pt(abs(t_statistic),df=dof))
                                
  }
  output <- data.frame(t_statistic,dof,p_value)
  return(output)
}

t_test_function(data=peng_AC,val_col="body_mass_g",group_col="species",var_equal=FALSE)
t.test(body_mass_g~species,data=peng_AC,var.equal=FALSE)
```

# 4 Investigating test size for an unpaired Student’s t-test

```{r}
num_trials<-1000
sample_size<-30
mu_0<-1
mu_1<-1
sigma_0<-3
sigma_1<-3
# alpha<-0.05
alpha<-seq(0,1,0.01)
set.seed(0) # set random seed for reproducibility

compute_type_1_error <- function(alpha){
  single_alpha_test_size_simulation_df<-data.frame(trial=seq(num_trials))%>%
    mutate(sample_0=map(.x=trial,.f=~rnorm(n=sample_size,mean=mu_0,sd=sigma_0)),
           sample_1=map(.x=trial,.f=~rnorm(n=sample_size,mean=mu_1,sd=sigma_1))) %>%
  # generate random Gaussian samples
    mutate(p_value=pmap(.l=list(trial,sample_0,sample_1),
                      .f=~t.test(..2,..3,var.equal = TRUE)$p.value)) %>%
  # generate p values
    mutate(type_1_error=p_value<alpha)
  
  t1e <- single_alpha_test_size_simulation_df%>%
    pull(type_1_error)%>%
    mean() # estimate of coverage probability
  
  return(t1e)
}

t1e <- map_dbl(.x=alpha,.f=~varying_alpha(.x))
alpha_t1e <- data.frame(alpha,t1e)%>%
  ggplot(aes(x=alpha,y=t1e))+
  geom_line()
alpha_t1e
```

# 5 The power of an unpaired t-test

```{r}
num_trials<-10000
n_0<-30
n_1<-30
mu_0<-3
mu_1<-4
sigma_0<-2
sigma_1<-2
alpha<-0.05
set.seed(0) # set random seed for reproducibility
data.frame(trial=seq(num_trials))%>%
mutate(sample_0=map(.x=trial,.f=~rnorm(n=n_0,mean=mu_0,sd=sigma_0)),
sample_1=map(.x=trial,.f=~rnorm(n=n_1,mean=mu_1,sd=sigma_1)))%>%
# generate random Gaussian samples
mutate(p_value=pmap(.l=list(trial,sample_0,sample_1),
.f=~t.test(..2,..3,var.equal = TRUE)$p.value))%>%
# generate p values
mutate(reject_null=p_value<alpha)%>%
pull(reject_null)%>%
mean() # estimate of coverage probability
```
































